---
title: "Coursera Capstone - Week 2"
date: 2016-07-02
author: suttonbm
layout: post
categories:
  - projects
tags:
  - coursera
  - data.science
  - R
project: datasciencecoursera.capstone
excerpt: >
  Creating a Simple Predictive Model
---

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

require(knitcitations)
bib.file <- "capstone-citations.bibtex"
bib <- read.bibtex(bib.file)
```

There are two tasks for week 2 of the SwiftKey Natural Language Processing Capstone for the Coursera Data Science specialization:

  * Exploratory Data Analysis
  * Simple Language Modeling

In this post, I'll cover my exploratory data analysis, and will follow up on the simple model construction in the [next post]({{ base.url }}/2016/07/nlp_simple_model/)

For more information on how I got to this point, refer to earlier posts about [NLP]({{ base.url }}/2016/06/nlp_intro/), [sampling data]({{ base.url }}/2016/06/sampling_data/), and [cleaning/tokenizing data]({{ base.url }}/2016/06/tokenize_clean_data/).

The data for this project are located at: [Dataset](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip)

### Introduction



### References
```{r echo=FALSE, results="asis"}
bibliography("html")
```
